Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-07-04 ms.tex
full readthrough up to monte carlo section

The solution of the radiative transfer equation for resonant scattering of Lyman $\alpha$ photons for a uniform sphere of constant gas density is found in the limit of large line center optical depth. A monochromatic source of photons is assumed at the center of the sphere, and the intensity within the sphere and at the surface is found with the Eddington approximation for the angular dependence. The solution can be represented a sum of two terms: a semi-analytic solution of the homogeneous equation which is required for the mean intensity $J=0$ at the surface, and a solution of the homogeneous equation which enforces the correct, zero-ingoing intensity boundary conditions at the surface. It is the compact, analytic form as well as the piece accounting for the correct, frequency dependent boundary condition that is novel in this study. The analytic solution is compared to numerical solutions with the Monte-Carlo method, which are valid at arbitrary optical depth, and the deviations from exact solution are investigated.

@@

The need for accurate distributions from which to draw outgoing photon properties motivates this work; namely, an analytic solution to the steady-state problem of \lya resonant scattering in spherical geometry. In addition, a semi-analytic solution to time-dependent \lya diffusion through the same medium is explored to quantify how long the photons are expected to spend in the domain before escaping. These solutions will be shown to accurately represent the photon distributions in frequency and time-to-escape at the surface of a spherical volume of gas, highlighting their potential for use as a Monte Carlo accelerant in magnetohydrodynamic calculations of phenomena involving \lya resonance-line radiation.

@@

\citet{1973MNRAS.162...43H, 1974MNRAS.166..373H} demonstrated that resonance-line radiation transfer in low-density, optically-thick media can be described by the Poisson equation. However, the accuracy of their solution is limited by the treatment of the boundary conditions. In their eigenfunction expansion, the spatial and frequency components of the differential equation are separated, yet the spatial eigenvalues are defined as being explicitly frequency-dependent. The resulting solution for large optical depths would suggest the mean intensity at the boundary of the slab is zero, which is unphysical. A more detailed discussion of the boundary condition is included in Section \ref{sec:steadystate}. \citet{1990ApJ...350..216N} follows this approach in an attempt to extend the result to media of intermediate optical depth, including the effects of scattering in the Doppler core of the line, but again does not use the correct boundary condition. \citet{?} improved on this approach, producing a divergent, inhomogeneous solution to the problem in 1D slab geometry that includes the effect of photon absorption within the medium. \citet{2006ApJ...649...14D} generalize the problem to spherical geometry, which is the same physical set up used here. Following this, \citet{2020arXiv200509692L} presents closed-form solutions for the radiation field in spherical geometry, utilizing the ``core skipping'' method discussed in \citet{2015MNRAS.449.4336S} to handle photon frequency redistribution within the Doppler core. This technique involves drawing the atom's velocity from a truncated Gaussian distribution, forcing photons back into the wing when they scatter into the core. This is motivated by the assumption that all core scattering results in zero spatial diffusion until the photon is redistributed back out into the wing where the mean free path becomes much larger and it can diffuse in space again. However, it is not clear whether this prescription is fully independent of the choice of velocity distribution.

Radiation forces due to \lya transfer have been calculated in \citet{1976ApJ...208..286W} in plane-parallel geometry. However, these solutions are limited to optical depths below $2.5 \times 10^3$. In the grid cells of a radiation magnetohydrodynamic simulation that includes resonant scattering, for example, it would be necessary to calculate the radiation force on each cell face due to \lya at line center optical depths of up to 1 million. A discussion of Monte Carlo code acceleration for optical depths as large as these in a uniform, plane-parallel medium with no absorption is presented in  \citet{2002ApJ...567..922A,2015MNRAS.449.4336S}. In terms of the time dependence, \citet{1994ApJ...427..603R} introduce an ansatz for the Fokker-Planck diffusion equation in order to quantify the relaxation timescale necessary to reach a quasi-static solution. Notably, the full Voigt function is used for the line profile rather than restricting scattering to the damping wing only.

@@

Photons of frequency $\nu$ near line center frequency $\nu_0$ are considered. The Doppler width will be written $\Delta = \nu_0 v_{\rm th}/c$, where $v_{\rm th}=\sqrt{2k_{\rm B}T/m_{\rm H}}$ is the thermal speed of hydrogen atoms of mass $m_{\rm H}$ and temperature $T$, and $c$ is the speed of light. The photon frequency in Doppler units will be written $x = (\nu-\nu_0)/\Delta$. For upper-state de-excitation rate $\Gamma$, the ratio of natural to Doppler broadening is $a=\Gamma/(4\pi \Delta)$. 
The transfer equation is \citep{1986rpa..book.....R}
\be
\frac{1}{c} \frac{\partial I}{\partial t} + \vec{n} \cdot \grad I & =& - \left( \alpha_{\rm sc} + \alpha_{\rm abs} \right) I + (1-p) j_{\rm sc} + j_{\rm em}.
\label{eq:rteqn}
\ee
The scattering coefficient, or inverse mean free path to scattering, is 
\be
\alpha_{\rm sc} & = & n_{\rm sc}\, \frac{\pi e^2}{m_e c}\, f\, \frac{H(x,a)}{\sqrt{\pi} \Delta}
= k \phi   
\ee
where $n_{\rm sc}$ is the number density of scatterers, $e$ and $m_e$ are the charge and mass of the electron, $f$ is the oscillator strength of the transition, $H(x,a)$ is the Voigt function, $k = n_{\rm sc} (\pi e^2/m_e c) f$, and the Voigt line profile is $\phi = H(x,a)/(\sqrt{\pi} \Delta)$, which is normalized as $\int d\nu\, \phi(\nu) = 1$. The absorption coefficient $\alpha_{\rm abs}$, or inverse mean free path to true absorption, is a sum over number density of the absorber times absorption cross section. Once the incoming photon has promoted the electron to an excited state, the collisional de-excitation probability is $p$, and hence only a fraction $1-p$ of the excitations lead to re-emission of photons. Photon destruction will not be considered, so $p$ is set to 0.

\subsection{Derivation of the Steady-State Solution}
Now, consider a sphere of radius $R$ and uniform density, with line-center optical depth $\tau_0 = (k/\sqrt{\pi}\Delta)R$. After changing variables from $\nu$ to $\sigma$ using $d\sigma = \sqrt{2/3}d\nu/(\Delta^2 \phi)$ in Eq. \ref{eq:rteqn} and performing all angular integrals, we obtain Eq. \ref{eq:finaleqn} (see Appendix \ref{app:rteqn_derivation}) and set $p=\alpha_{\rm abs}=0$ to eliminate photon destruction terms. With a photon emission term given by Eq. \ref{eq:jem}, this gives the transfer equation
\be
\nabla^2 J + \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} & = & 
- \frac{ \sqrt{6} kL}{4\pi \Delta^2} \delta^3(\vec{x} - \vec{x}_s) \delta (\sigma - \sigma_{\rm s})
\label{eq:rt_no_destr}
\ee
where the form of the source term is in Eq. \ref{eq:jem_v2}. The boundary condition is that there is no incoming flux at the surface, which can be written
\citep{1986rpa..book.....R}
\be
J & = & \sqrt{3} H
\label{eq:bc}
\ee
at $r=R$. An analytic solution $J_d$ which is divergent at the center of the sphere and at the emission frequency may be found. However, it is not a good approximation to the true solution, as it is too large at $r=R$ by a factor of $J_{\rm d}(R,\sigma)/ H_{\rm d}(R,\sigma) \sim a\tau_0/x^2 \sim (a\tau_0)^{1/3} \gg 1$. This solution is included in Fig. \ref{fig:sol_mc_residual_0} for qualitative understanding, but will not be discussed further.

@@

The question is how to satisfy the surface boundary condition in Eq. \ref{eq:bc}. On this point, we note that the derivation in \citet{1973MNRAS.162...43H}, and numerous studies later studies, does not actually produce a valid solution of the transfer equation and boundary condition. The reason is as follows. In the notation of \citet{1973MNRAS.162...43H}, a separation of variables
$J(\tau,\sigma) = \theta(\tau) j(\sigma)$ in spatial variable $\tau$ and frequency variable $\sigma$ is attempted (their Equations 16 and 23). The solutions for the eigenfunctions $\theta(\tau)$ and $j(\sigma)$ then depend explicitly on the separation constant $\lambda$. In order to satisfy the boundary conditions, the separation constant must satisfy an eigenvalue equation of the form
\be
\lambda \tan(\lambda B) & = & \frac{3}{2} \phi,
\label{eq:evalue}
\ee
where $2B$ is the slab optical depth at line center, and $\phi$ is the line profile ($\Delta \phi(\nu)$ here). The key point is that the line profile depends on one of the coordinates, frequency, and this causes the eigenvalues of the separation constant to depend on frequency. But then the separation constant is not constant, and it does not satisfy Eq. \ref{eq:rt_no_destr} since the frequency derivatives will act on the separation ``constant", giving extra terms. To our knowledge the errors introduced by this ansatz have never been quantified.

@@

A different solution method is attempted here: namely, a continuous Fourier expansion in the frequency variable $\sigma$. The full solution is written
\be
J(r,\sigma) & = & J_0(r,\sigma) + J_{\rm bc}(r,\sigma)
\ee
and
\be \label{eq:totalflux}
H(r,\sigma) & = & H_0(r,\sigma) + H_{\rm bc}(r,\sigma).
\ee
Here, $J_0$ and $H_0$ are the solutions of the homogeneous equation, given in Equations \ref{eq:J0} and \ref{eq:H0}. The additional term $J_{\rm bc}$ must then be a solution of the homogeneous equation
\be \label{eq:diffeq}
\frac{\partial^2J}{\partial r^2} + \frac{2}{r} \frac{\partial J}{\partial r}
+ \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} &= & 0
\ee
with no delta function source term, and it must allow the boundary conditions to be satisfied at the surface. Since $J_0(R,
\nu)=0$, the surface boundary condition becomes
\be
J_{\rm bc}(R,\sigma) - \sqrt{3} H_{\rm bc}(R,\sigma) & = & 
J_{0}(R,\sigma) + \frac{1}{\sqrt{3}k\phi} \frac{\partial J_{0}(R,\sigma)}{\partial r} = 
\sqrt{3} H_0(R,\sigma).
\label{eq:bc2}
\ee
Plugging in a frequency dependence $J \propto e^{is\sigma}$, for ``wavenumber" $s$, gives the equation for modified spherical Bessel functions of the first kind, $i_0(z)=\sinh(z)/z$. The solution can then be represented as
\be
J_{\rm bc}(r,\sigma) & = & 
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\frac{i_0(krs/\Delta)}{i_0(kRs/\Delta)},
\label{eq:Jbc}
\ee
where $A(s)$ is the Fourier amplitude. Plugging Eq. \ref{eq:Jbc} into Eq. \ref{eq:bc2} leads to the following equation for the Fourier amplitudes,
\be
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s)
\left[ 1 + \left( \frac{s}{\sqrt{3} \Delta \phi} \right) \left( \frac{i_0^\prime(kRs/\Delta)}{i_0(kRs/\Delta)} \right) \right]
& = & \sqrt{3} H_0(R,\sigma)
\label{eq:bc3}
\ee
where $H_0(R,\sigma)$ is given by Eq. \ref{eq:H0surf}. Discretization of Eq. \ref{eq:bc3} for frequency variables $\sigma_i$ and wavenumbers $s_j$
leads to a set of coupled linear equations for the $A(s_j)$. We use equally-spaced points $\Delta \sigma = 2\sigma_{\rm max}/(N-1)$ and $\Delta s = 2\pi/(N\Delta \sigma)$, where $N$ is the number of points for each grid and solve the resulting matrix equation. The maximum frequency is set as $\sigma_{\rm max} = {\rm constant} \times \tau_0$, for a large enough constant that  the end of the frequency grid is at such small intensities that it does not affect the solution except close to the boundaries. The number of points was increased until the solution was well-resolved near line center, and only became inaccurate near the boundaries. Given the Fourier amplitudes, $J_{\rm bc}$ is computed using Eq. \ref{eq:Jbc}, and the flux is given by
\be
H_{\rm bc}(r,\sigma) & = & \left( \frac{-1}{3k\phi} \right)
\frac{\partial J_{\rm bc}(r,\sigma)}{\partial r}
= \left( \frac{-1}{3k\phi} \right)
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\left( \frac{ks}{\Delta} \right) 
\left( \frac{i_0^\prime(krs/\Delta)}{i_0(kRs/\Delta)} \right).
\label{eq:Hbc}
\ee
The Bessel functions are finite at the center and rise steeply toward the surface when $kRs/\Delta \gg 1$.

\subsection{Scaling with Line Center Optical Depth}
To examine the relative size of the corrective term $H_{\rm bc}$ as compared to $H_0$, we consider the case where the optical depth is large and thus the mean intensity will be considerably larger than the surface flux. Starting from Eq. \ref{eq:bc2}, we assume $H_{\rm bc} \ll J_{\rm bc}$ such that
\be
J_{\rm bc}(R,\sigma) & \approx & \sqrt{3} H_0(R,\sigma).
\label{eq:bc4}
\ee
The Fourier transform for $J_{\rm bc}$ at the surface becomes
\be
J_{\rm bc}(R, \sigma) \equiv \int_{-\infty}^{\infty} \frac{ds}{2\pi} e^{is\sigma} A(s)
\ee
with Fourier amplitudes
\be \label{eq:amps}
A(s) = \sqrt{3} \int_{-\infty}^{\infty} d\sigma e^{-is\sigma} H_0(R, \sigma).
\ee
Plugging the expression for $H_0$ in Eq. \ref{eq:H0surf} into Eq. \ref{eq:amps} yields an expression for the Fourier amplitudes
\be \label{eq:plugged_amps}
A(s) = \sqrt{3}\frac{1}{3k}\frac{\sqrt{6}}{32\pi}\frac{L}{\Delta R^3} \int_{-\infty}^\infty d\sigma e^{-is\sigma}\frac{1}{\phi(\sigma)}\frac{1}{1+\cosh \left[\frac{\pi \Delta}{kR} |\sigma - \sigma_i|\right]}
\ee
with
\be
\phi(\sigma) = \left(\frac{2a}{27\pi}\right)^{1/3}\frac{1}{\Delta |\sigma|^{2/3}}.
\ee
With the line center optical depth defined as
\be \label{eq:tau0}
\tau_0 = \frac{kR}{\sqrt{\pi}\Delta}
\ee
we scale the variables $\sigma$ and $s$ by $\tau_0$, now using characteristic photon frequency $\bar \sigma = \sigma / \tau_0$ and characteristic wavenumber $\bar s = s \tau_0$. Eq. \ref{eq:plugged_amps} becomes
\be
A(\bar s) = \frac{\sqrt{2}}{32\pi}\left(\frac{27\pi}{2}\right)^{1/3}\frac{L}{kR^3}\frac{\tau_0^{5/3}}{a^{1/3}}\int_{-\infty}^\infty d\bar\sigma e^{-i\bar s \bar\sigma} |\bar\sigma^{2/3}|\frac{1}{1+\cosh \left[\sqrt{\pi}|\bar\sigma - \bar\sigma_i|\right]}.
\ee
These coefficients are used to evaluate Eq. \ref{eq:Hbc} at the surface of the sphere. Dividing $H_{\rm bc}$ by $H_0$ (Eq. \ref{eq:H0surf}) yields an overall proportionality
\be \label{eq:hbc_scaling}
\frac{H_{\rm bc}(R, \sigma)}{H_0(R, \sigma)} \propto \frac{1}{(a\tau_0)^{1/3}}.
\ee
Thus, the fractional importance of the corrective frequency-dependent boundary condition term scales as the inverse cube root of the damping parameter $a$ times the line-center optical depth $\tau_0$. At very large optical depths, it is expected that the corrective term will be small, but will become increasingly important in less optically thick media.

>>>>

The solution of the radiative transfer equation for resonant scattering of Lyman $\alpha$ photons for a uniform sphere of constant gas density is found in the limit of large line center optical depth. A monochromatic source of photons is assumed at the center of the sphere, and the intensity within the sphere and at the surface is found with the Eddington approximation for the angular dependence. The solution can be represented a sum of two terms: a semi-analytic solution of the homogeneous equation which is required for the mean intensity $J=0$ at the surface, and a solution of the homogeneous equation which enforces the correct, zero-ingoing intensity boundary conditions at the surface. It is the compact analytic form as well as the piece accounting for the correct, frequency-dependent boundary condition that is novel in this study. The analytic solution is compared to numerical solutions with the Monte-Carlo method, which are valid at arbitrary optical depth. Discrepancies in the spectra of previous work are quantified, and it is shown that the surface boundary condition derived here corrects these discrepancies. Separately, the time-dependence of the homogeneous equation is calculated using an eigenfunction expansion and solved numerically to characterize the time-to-escape of photons initialized at the center of the sphere, including the effects of frequency redistribution. This distribution can be sampled to accelerate Monte Carlo radiation transfer calculations at very large optical depths, obtaining intensities and radiation forces on cell faces at a fraction of the computational cost.

@@

The need for accurate distributions from which to draw outgoing photons' properties motivates this work; namely, an analytic solution to the steady-state problem of \lya resonant scattering in spherical geometry. In addition, a semi-analytic solution to time-dependent \lya diffusion through the same medium is explored to quantify how long the photons are expected to spend in the domain before escaping. These solutions will be shown to accurately represent the photon distributions in frequency and time-to-escape at the surface of a spherical volume of gas, highlighting their potential for use as a Monte Carlo accelerant in magnetohydrodynamic calculations of phenomena involving \lya resonance-line radiation.

@@

\citet{1973MNRAS.162...43H, 1974MNRAS.166..373H} demonstrated that resonance-line radiation transfer in low-density, optically-thick media can be described by the Poisson equation. However, the accuracy of their solution is limited by the treatment of the boundary conditions. In their eigenfunction expansion, the spatial and frequency components of the differential equation are separated, yet the spatial eigenvalues are defined as being explicitly frequency-dependent. The resulting solution is therefore invalid on the basis that the differential equation and boundary conditions cannot simultaneously be satisfied. A more detailed discussion of the boundary condition is included in Section \ref{sec:steadystate}. \citet{1990ApJ...350..216N} follows this approach in an attempt to extend the result to media of intermediate optical depth, including the effects of scattering in the Doppler core of the line, but again does not use the correct boundary condition. \citet{?} improved on this approach, producing a divergent, inhomogeneous solution to the problem in 1D slab geometry that includes the effect of photon absorption within the medium. \citet{2006ApJ...649...14D} generalize the problem to spherical geometry, which is the same physical set up used here. Following this, \citet{2020arXiv200509692L} presents closed-form solutions for the radiation field in spherical geometry, utilizing the ``core skipping'' method discussed in \citet{2015MNRAS.449.4336S} to handle photon frequency redistribution within the Doppler core. This technique involves drawing the atom's velocity from a truncated Gaussian distribution, forcing photons back into the wing when they scatter into the core. This is motivated by the assumption that all core scattering results in zero spatial diffusion until the photon is redistributed back out into the wing where the mean free path becomes much larger and it can diffuse in space again. However, it is not clear whether this prescription is fully independent of the choice of velocity distribution.

Radiation forces due to \lya transfer have been calculated in \citet{1976ApJ...208..286W} in plane-parallel geometry. However, these solutions are limited to optical depths below $2.5 \times 10^3$. In the grid cells of a radiation hydrodynamic simulation, it would be necessary to calculate the radiation force on each cell face due to \lya at line center optical depths of up to 1 million or more. A discussion of Monte Carlo code acceleration for optical depths as large as these in a uniform, plane-parallel medium with no absorption is presented in  \citet{2002ApJ...567..922A,2015MNRAS.449.4336S}. In terms of the time dependence, \citet{1994ApJ...427..603R} introduce an ansatz for the Fokker-Planck diffusion equation in order to quantify the relaxation timescale necessary to reach a quasi-static solution. Notably, the full Voigt function is used for the line profile rather than restricting scattering to the damping wing only.

@@

Photons of frequency $\nu$ near line center frequency $\nu_0$ are considered. The Doppler width will be written $\Delta = \nu_0 v_{\rm th}/c$, where $v_{\rm th}=\sqrt{2k_{\rm B}T/m_{\rm H}}$ is the thermal speed of hydrogen atoms of mass $m_{\rm H}$ and temperature $T$, and $c$ is the speed of light. The photon frequency in Doppler units will be written $x = (\nu-\nu_0)/\Delta$. For upper-state de-excitation rate $\Gamma$, the ratio of natural to Doppler broadening is $a=\Gamma/(4\pi \Delta)$. 
The transfer equation is \citep{1986rpa..book.....R}
\be
\frac{1}{c} \frac{\partial I}{\partial t} + \vec{n} \cdot \grad I & =& - \left( \alpha_{\rm sc} + \alpha_{\rm abs} \right) I + (1-p) j_{\rm sc} + j_{\rm em}.
\label{eq:rteqn}
\ee
The scattering coefficient, or inverse mean free path to scattering, is 
\be
\alpha_{\rm sc} & = & n_{\rm sc}\, \frac{\pi e^2}{m_e c}\, f\, \frac{H(x,a)}{\sqrt{\pi} \Delta}
= k \phi   
\ee
where $n_{\rm sc}$ is the number density of scatterers, $e$ and $m_e$ are the charge and mass of the electron, $f$ is the oscillator strength of the transition, $H(x,a)$ is the Voigt function, $k = n_{\rm sc} (\pi e^2/m_e c) f$, and the Voigt line profile is $\phi = H(x,a)/(\sqrt{\pi} \Delta)$, which is normalized as $\int d\nu\, \phi(\nu) = 1$. The absorption coefficient $\alpha_{\rm abs}$, or inverse mean free path to true absorption, is a sum over number density of the absorber times absorption cross section. Once the incoming photon has promoted the electron to an excited state, the collisional de-excitation probability is $p$, and hence only a fraction $1-p$ of the excitations lead to re-emission of photons. Photon destruction will not be considered, so $p$ is set to 0 in this work.

\subsection{Derivation of the Steady-State Solution}
Now, consider a sphere of radius $R$ and uniform density, with line-center optical depth $\tau_0 = (k/\sqrt{\pi}\Delta)R$. After changing variables from $\nu$ to $\sigma$ using $d\sigma = \sqrt{2/3}d\nu/(\Delta^2 \phi)$ in Eq. \ref{eq:rteqn} and performing all angular integrals, we obtain Eq. \ref{app:rteqn_derivation}\ref{eq:finaleqn} (see Appendix \ref{app:rteqn_derivation}) and set $p=\alpha_{\rm abs}=0$ to eliminate photon destruction terms. With a photon emission term given by Eq. \ref{app:rteqn_derivation}\ref{eq:jem}, this gives the transfer equation
\be
\nabla^2 J + \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} & = & 
- \frac{ \sqrt{6} kL}{4\pi \Delta^2} \delta^3(\vec{x} - \vec{x}_s) \delta (\sigma - \sigma_{\rm s})
\label{eq:rt_no_destr}
\ee
where the form of the source term is in Eq. \ref{app:rteqn_derivation}\ref{eq:jem_v2}. The boundary condition is that there is no incoming flux at the surface, which can be written
\citep{1986rpa..book.....R}
\be
J & = & \sqrt{3} H
\label{eq:bc}
\ee
at $r=R$. An analytic solution $J_d$ which is divergent at the center of the sphere and at the emission frequency may be found. However, it is not a good approximation to the true solution, as it is too large at $r=R$ by a factor of $J_{\rm d}(R,\sigma)/ H_{\rm d}(R,\sigma) \sim a\tau_0/x^2 \sim (a\tau_0)^{1/3} \gg 1$. This solution is included in Fig. \ref{fig:sol_mc_residual_0} for qualitative understanding, but will not be discussed further.

@@

The question is how to satisfy the surface boundary condition in Eq. \ref{eq:bc}. On this point, we note that the derivation in \citet{1973MNRAS.162...43H}, and numerous studies later studies, does not actually produce a valid solution of the transfer equation and boundary condition. The reason is as follows. In the notation of \citet{1973MNRAS.162...43H}, a separation of variables
$J(\tau,\sigma) = \theta(\tau) j(\sigma)$ in spatial variable $\tau$ and frequency variable $\sigma$ is attempted (their Equations 16 and 23). The solutions for the eigenfunctions $\theta(\tau)$ and $j(\sigma)$ then depend explicitly on the separation constant $\lambda$. In order to satisfy the boundary conditions, the separation constant must satisfy an eigenvalue equation of the form
\be
\lambda \tan(\lambda B) & = & \frac{3}{2} \phi,
\label{eq:evalue}
\ee
where $2B$ is the slab optical depth at line center, and $\phi$ is the line profile ($\phi(\nu)\Delta$ here). The key point is that the line profile depends on one of the coordinates, frequency, and this causes the eigenvalues of the separation constant to depend on frequency. But then the separation constant is not constant, and it does not satisfy Eq. \ref{eq:rt_no_destr} since the frequency derivatives will act on the separation ``constant", giving extra terms. To our knowledge the errors introduced by this ansatz have never been quantified.

@@

A different solution method is attempted here: namely, a continuous Fourier expansion in the frequency variable $\sigma$. The full solution is written
\be
J(r,\sigma) & = & J_0(r,\sigma) + J_{\rm bc}(r,\sigma)
\ee
and
\be \label{eq:totalflux}
H(r,\sigma) & = & H_0(r,\sigma) + H_{\rm bc}(r,\sigma).
\ee
Here, $J_0$ and $H_0$ are the solutions of the homogeneous equation, given in Equations \ref{eq:J0} and \ref{eq:H0}. The additional term $J_{\rm bc}$ must then be a solution of the homogeneous equation
\be \label{eq:diffeq}
\frac{\partial^2J}{\partial r^2} + \frac{2}{r} \frac{\partial J}{\partial r}
+ \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2} &= & 0
\ee
with no delta function source term, and it must allow the boundary conditions to be satisfied at the surface. Since $J_0(R,
\sigma)=0$, the surface boundary condition becomes
\be
J_{\rm bc}(R,\sigma) - \sqrt{3} H_{\rm bc}(R,\sigma) & = & 
J_{0}(R,\sigma) + \frac{1}{\sqrt{3}k\phi} \frac{\partial J_{0}(R,\sigma)}{\partial r} = 
\sqrt{3} H_0(R,\sigma).
\label{eq:bc2}
\ee
Plugging in a frequency dependence $J \propto e^{is\sigma}$, for ``wavenumber" $s$, gives the equation for modified spherical Bessel functions of the first kind, $i_0(z)=\sinh(z)/z$. The solution can then be represented as
\be
J_{\rm bc}(r,\sigma) & = & 
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\frac{i_0(krs/\Delta)}{i_0(kRs/\Delta)},
\label{eq:Jbc}
\ee
where $A(s)$ is the Fourier amplitude. Plugging Eq. \ref{eq:Jbc} into Eq. \ref{eq:bc2} leads to the following equation for the Fourier amplitudes,
\be
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s)
\left[ 1 + \left( \frac{s}{\sqrt{3} \Delta \phi} \right) \left( \frac{i_0^\prime(kRs/\Delta)}{i_0(kRs/\Delta)} \right) \right]
& = & \sqrt{3} H_0(R,\sigma)
\label{eq:bc3}
\ee
where $H_0(R,\sigma)$ is given by Eq. \ref{eq:H0surf}. Discretization of Eq. \ref{eq:bc3} for frequency variables $\sigma_i$ and wavenumbers $s_j$
leads to a set of coupled linear equations for the $A(s_j)$. We use equally-spaced points $\Delta \sigma = 2\sigma_{\rm max}/(N-1)$ and $\Delta s = 2\pi/(N\Delta \sigma)$, where $N$ is the number of points for each grid and solve the resulting matrix equation. The maximum frequency is set as $\sigma_{\rm max} = {\rm constant} \times \tau_0$, for a large enough constant that  the end of the frequency grid is at such small intensities that it does not affect the solution except close to the boundaries. The number of points was increased until the solution was well-resolved near line center, and only became inaccurate near the boundaries. Given the Fourier amplitudes, $J_{\rm bc}$ is computed using Eq. \ref{eq:Jbc}, and the flux is given by
\be
H_{\rm bc}(r,\sigma) & = & \left( \frac{-1}{3k\phi} \right)
\frac{\partial J_{\rm bc}(r,\sigma)}{\partial r}
= \left( \frac{-1}{3k\phi} \right)
\int_{-\infty}^\infty \frac{ds}{2\pi} e^{is\sigma} A(s) 
\left( \frac{ks}{\Delta} \right) 
\left( \frac{i_0^\prime(krs/\Delta)}{i_0(kRs/\Delta)} \right).
\label{eq:Hbc}
\ee
The Bessel functions are finite at the center and rise steeply toward the surface when $kRs/\Delta \gg 1$.

\subsection{Scaling with Line Center Optical Depth}
To examine the relative size of the corrective term $H_{\rm bc}$ as compared to $H_0$, we consider the case where the optical depth is large and thus the mean intensity will be considerably larger than the surface flux. Starting from Eq. \ref{eq:bc2}, we assume $H_{\rm bc} \ll J_{\rm bc}$ such that
\be
J_{\rm bc}(R,\sigma) & \approx & \sqrt{3} H_0(R,\sigma).
\label{eq:bc4}
\ee
The Fourier transform for $J_{\rm bc}$ at the surface becomes
\be
J_{\rm bc}(R, \sigma) \equiv \int_{-\infty}^{\infty} \frac{ds}{2\pi} e^{is\sigma} A(s)
\ee
with Fourier amplitudes
\be \label{eq:amps}
A(s) = \sqrt{3} \int_{-\infty}^{\infty} d\sigma e^{-is\sigma} H_0(R, \sigma).
\ee
Plugging the expression for $H_0$ in Eq. \ref{eq:H0surf} into Eq. \ref{eq:amps} yields an expression for the Fourier amplitudes
\be \label{eq:plugged_amps}
A(s) = \sqrt{3} \int_{-\infty}^\infty d\sigma e^{-is\sigma}\left(\frac{1}{3k\phi(\sigma)}\frac{\sqrt{6}}{32\pi}\frac{L}{\Delta R^3}\frac{1}{1+\cosh \left[\frac{\pi \Delta}{kR} |\sigma - \sigma_i|\right]}\right)
\ee
with
\be
\phi(\sigma) = \left(\frac{2a}{27\pi}\right)^{1/3}\frac{1}{\Delta |\sigma|^{2/3}}.
\ee
With the line center optical depth defined as
\be \label{eq:tau0}
\tau_0 = \frac{kR}{\sqrt{\pi}\Delta}
\ee
we scale the variables $\sigma$ and $s$ by $\tau_0$, now using characteristic photon frequency $\bar \sigma = \sigma / \tau_0$ and characteristic wavenumber $\bar s = s \tau_0$. Eq. \ref{eq:plugged_amps} becomes
\be
A(\bar s) = \frac{\sqrt{2}}{32\pi}\left(\frac{27\pi}{2}\right)^{1/3}\frac{L}{kR^3}\frac{\tau_0^{5/3}}{a^{1/3}}\int_{-\infty}^\infty d\bar\sigma e^{-i\bar s \bar\sigma} |\bar\sigma^{2/3}|\frac{1}{1+\cosh \left[\sqrt{\pi}|\bar\sigma - \bar\sigma_i|\right]}.
\ee
These coefficients are used to evaluate the boundary condition's contribution to the flux (Eq. \ref{eq:Hbc}) at the surface of the sphere. Dividing this additional flux $H_{\rm bc}$ by the contribution from the $J=0$ solution, $H_0$ (Eq. \ref{eq:H0surf}), yields an overall proportionality
\be \label{eq:hbc_scaling}
\frac{H_{\rm bc}(R, \sigma)}{H_0(R, \sigma)} \propto \frac{1}{(a\tau_0)^{1/3}}.
\ee
Thus, the fractional importance of the corrective frequency-dependent boundary condition term scales as the inverse cube root of the damping parameter $a$ times the line-center optical depth $\tau_0$. At very large optical depths, it is expected that the corrective term will be small, but will become increasingly important in less optically thick media.

By replacing the differential operators in Eq. \ref{eq:rt_no_destr} with characteristic scales $R$ and $\sigma_0$ and equating the terms on the left-hand side of this equation, a dimensionless scaling relation between frequency and space emerges.
\be
\frac{J}{R^2} &=& \left(\frac{k}{\Delta}\right)^2 \frac{J}{\sigma_0^2}
\ee
Now, using the line center optical depth (Eq. \ref{eq:tau0}) and the approximation $\sigma(x)=x^3\sqrt{2\pi/27}/a$ introduced in \cite{1973MNRAS.162...43H} for large $x$, we have
\be
\frac{\sqrt{2\pi/27}}{a}x_0^3 = \frac{kR}{\Delta} = \sqrt{\pi} \tau_0.
\ee
Simplifying, we obtain the proportionality
\be \label{eq:tau_peak_scaling}
x_0 &\propto& \left(a\tau_0\right)^{1/3}.
\ee
This predicts that the characteristic photon frequency in Doppler widths should scale as the cube root of the parameter $a$ multiplied by the line center optical depth $\tau_0$, indicating the location of the spectral peaks on either side of line center --- the locations where the photons most readily escape. Since our derivations were performed assuming a Lorentzian line profile (see Appendix \ref{app:rteqn_derivation}), our solutions are technically only valid in the line wing, where the Doppler component of the line profile is small. Since our choice of the damping parameter $a$ is constant, the solutions of the transfer equation will perform best when the peaks of the spectral energy distribution lie well outside of the Doppler core, e.g., for large optical depths. The boundary of the Doppler core is approximately where the Gaussian and Lorentzian pieces are equal:
\be
\frac{e^{-x^2}}{\sqrt{\pi}} = \frac{a}{\pi x^2}
\ee
For \lya and a hydrogen gas of temperature $T=10^4$ K, the damping parameter $a = \Gamma / (4\pi\Delta)$ is equal to $4.72\times 10^{-4}$ in CGS units. The Doppler core is then significant over the range $-4.6 < x < 4.6$. Comparing with the locations of the spectral peaks at various optical depths using Eq. \ref{eq:tau_peak_scaling}, we find that $x_0=$3.6, 7.8, 16.8 for $\tau_0=10^5, 10^6,$ and $10^7$, respectively. Therefore, solutions derived assuming a Lorentzian line profile $\phi(x) \approx a/(\pi x^2)$ are expected to fail for lower line center optical depths where the peak of the spectrum lies near the core and improve for larger optical depths where the spectral peak is well outside of the line core.
\subsection{Comparison to Monte Carlo}
